---
title: "Project_part1"
author: "Jingzhi Zhou"
date: "11/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
setwd("C:/Users/11638/Dropbox/My PC (LAPTOP-5N5V9K0V)/Desktop/535/ISE 535/Project/DATA")
library(readr)
library(dplyr)
library(tidyverse)
```

# Load Data
```{r}
breast_cancer <- read_csv("wdbc.data", 
                          col_names = FALSE, col_types = cols(X1 = col_character()))

names1 <- c('id_number', 'diagnosis', 'radius_mean', 'texture_mean', 'perimeter_mean', 'area_mean', 'smoothness_mean', 'compactness_mean', 'concavity_mean','concave_points_mean', 
            'symmetry_mean', 'fractal_dimension_mean', 'radius_se', 'texture_se', 'perimeter_se', 
            'area_se', 'smoothness_se', 'compactness_se', 'concavity_se', 'concave_points_se', 'symmetry_se', 'fractal_dimension_se', 'radius_worst', 'texture_worst', 'perimeter_worst', 'area_worst', 'smoothness_worst', 'compactness_worst', 'concavity_worst', 'concave_points_worst', 'symmetry_worst', 'fractal_dimension_worst')

colnames (breast_cancer) <- names1
```

# Creating two subsets
```{r}
M_breast_cancer = breast_cancer[breast_cancer$diagnosis=="M",]
B_breast_cancer = breast_cancer[breast_cancer$diagnosis=="B",]
```

# Summary and Correlation
```{r}
pairs(M_breast_cancer[,3:12])
summary(M_breast_cancer[,3:12])
```
```{r}
library(corrplot)
par(mfrow=c(1,2))
X1 = cor(M_breast_cancer[,3:12])
corrplot(X1, method="color", main = "Mean: Malignant Class")

X2 = cor(B_breast_cancer[,3:12])
corrplot(X2, method="color", main = "Mean: Benign Class")

X3 = cor(M_breast_cancer[,13:22])
corrplot(X3, method="color", main = "SE: Malignant Class")

X4 = cor(B_breast_cancer[,13:22])
corrplot(X4, method="color", main = "SE: Benign Class")

X5 = cor(M_breast_cancer[,23:32])
corrplot(X5, method="color", main = "Worst: Malignant Class")

X6 = cor(B_breast_cancer[,23:32])
corrplot(X6, method="color", main = "Worst: Benign Class")
```

```{r}
X1 = cor(B_breast_cancer[,3:12])
corrplot(X1, method="color")
X2 = cor(B_breast_cancer[,13:22])
corrplot(X2, method="color")
X3 = cor(B_breast_cancer[,23:32])
corrplot(X3, method="color")
```

# Boxplots
#of means
```{r}
par(mfrow=c(2,5))
boxplot(radius_mean~diagnosis, data=breast_cancer)
boxplot(texture_mean~diagnosis,data=breast_cancer)
boxplot(perimeter_mean~diagnosis,data=breast_cancer)
boxplot(area_mean~diagnosis, data=breast_cancer)
boxplot(smoothness_mean~diagnosis,data=breast_cancer)
boxplot(compactness_mean~diagnosis,data=breast_cancer)
boxplot(concavity_mean~diagnosis,data=breast_cancer)
boxplot(concave_points_mean~diagnosis,data=breast_cancer)
boxplot(symmetry_mean~diagnosis,data=breast_cancer)
boxplot(fractal_dimension_mean~diagnosis,data=breast_cancer)
```
#of se
```{r}
par(mfrow=c(2,5))
boxplot(radius_se~diagnosis, data=breast_cancer)
boxplot(texture_se~diagnosis,data=breast_cancer)
boxplot(perimeter_se~diagnosis,data=breast_cancer)
boxplot(area_se~diagnosis, data=breast_cancer)
boxplot(smoothness_se~diagnosis,data=breast_cancer)
boxplot(compactness_se~diagnosis,data=breast_cancer)
boxplot(concavity_se~diagnosis,data=breast_cancer)
boxplot(concave_points_se~diagnosis,data=breast_cancer)
boxplot(symmetry_se~diagnosis,data=breast_cancer)
boxplot(fractal_dimension_se~diagnosis,data=breast_cancer)
```
#of worst
```{r}
par(mfrow=c(2,5))
boxplot(radius_worst~diagnosis, data=breast_cancer)
boxplot(texture_worst~diagnosis,data=breast_cancer)
boxplot(perimeter_worst~diagnosis,data=breast_cancer)
boxplot(area_worst~diagnosis, data=breast_cancer)
boxplot(smoothness_worst~diagnosis,data=breast_cancer)
boxplot(compactness_worst~diagnosis,data=breast_cancer)
boxplot(concavity_worst~diagnosis,data=breast_cancer)
boxplot(concave_points_worst~diagnosis,data=breast_cancer)
boxplot(symmetry_worst~diagnosis,data=breast_cancer)
boxplot(fractal_dimension_worst~diagnosis,data=breast_cancer)
```

# Histograms of whole dataset
```{r}
par(mfrow=c(2,3))      
hist(breast_cancer$radius_mean)
hist(breast_cancer$texture_mean)
hist(breast_cancer$perimeter_mean)
hist(breast_cancer$area_mean)
hist(breast_cancer$smoothness_mean)
hist(breast_cancer$compactness_mean)
hist(breast_cancer$concavity_mean)
hist(breast_cancer$concave_points_mean)
hist(breast_cancer$symmetry_mean)
hist(breast_cancer$fractal_dimension_mean)
```

# Histograms of Diagnosis M
```{r}
par(mfrow=c(2,3))
hist(M_breast_cancer$radius_mean)
hist(M_breast_cancer$texture_mean)
hist(M_breast_cancer$perimeter_mean)
hist(M_breast_cancer$area_mean)
hist(M_breast_cancer$smoothness_mean)
hist(M_breast_cancer$compactness_mean)
hist(M_breast_cancer$concavity_mean)
hist(M_breast_cancer$concave_points_mean)
hist(M_breast_cancer$symmetry_mean)
hist(M_breast_cancer$fractal_dimension_mean)
```

# Histograms of Diagnosis B
```{r}
par(mfrow=c(2,3))
hist(B_breast_cancer$radius_mean)
hist(B_breast_cancer$texture_mean)
hist(B_breast_cancer$perimeter_mean)
hist(B_breast_cancer$area_mean)
hist(B_breast_cancer$smoothness_mean)
hist(B_breast_cancer$compactness_mean)
hist(B_breast_cancer$concavity_mean)
hist(B_breast_cancer$concave_points_mean)
hist(B_breast_cancer$symmetry_mean)
hist(B_breast_cancer$fractal_dimension_mean)
```
#SE
```{r}
par(mfrow=c(2,5))
hist(M_breast_cancer$radius_se)
hist(M_breast_cancer$texture_se)
hist(M_breast_cancer$perimeter_se)
hist(M_breast_cancer$area_se)
hist(M_breast_cancer$smoothness_se)
hist(M_breast_cancer$compactness_se)
hist(M_breast_cancer$concavity_se)
hist(M_breast_cancer$concave_points_se)
hist(M_breast_cancer$symmetry_se)
hist(M_breast_cancer$fractal_dimension_se)
```

```{r}
par(mfrow=c(2,5))
hist(B_breast_cancer$radius_se)
hist(B_breast_cancer$texture_se)
hist(B_breast_cancer$perimeter_se)
hist(B_breast_cancer$area_se)
hist(B_breast_cancer$smoothness_se)
hist(B_breast_cancer$compactness_se)
hist(B_breast_cancer$concavity_se)
hist(B_breast_cancer$concave_points_se)
hist(B_breast_cancer$symmetry_se)
hist(B_breast_cancer$fractal_dimension_se)
```

#Worst
```{r}
par(mfrow=c(2,5))
hist(M_breast_cancer$radius_worst)
hist(M_breast_cancer$texture_worst)
hist(M_breast_cancer$perimeter_worst)
hist(M_breast_cancer$area_worst)
hist(M_breast_cancer$smoothness_worst)
hist(M_breast_cancer$compactness_worst)
hist(M_breast_cancer$concavity_worst)
hist(M_breast_cancer$concave_points_worst)
hist(M_breast_cancer$symmetry_worst)
hist(M_breast_cancer$fractal_dimension_worst)
```

```{r}
par(mfrow=c(2,5))
hist(B_breast_cancer$radius_worst)
hist(B_breast_cancer$texture_worst)
hist(B_breast_cancer$perimeter_worst)
hist(B_breast_cancer$area_worst)
hist(B_breast_cancer$smoothness_worst)
hist(B_breast_cancer$compactness_worst)
hist(B_breast_cancer$concavity_worst)
hist(B_breast_cancer$concave_points_worst)
hist(B_breast_cancer$symmetry_worst)
hist(B_breast_cancer$fractal_dimension_worst)
```

# Multiple Logistic Model

#3-12,13-22,23-32

# drop SE
```{r}
breast_cancer1 <- breast_cancer[-c(13:22)]
```

# new column isM: if diagnosis is malignant, column value is 1, otherwise it is 0.
```{r}
breast_cancer1$isM <- as.factor(ifelse(breast_cancer1$diagnosis=="M",1,0))
breast_cancer1$isM <- as.numeric(as.character(breast_cancer1$isM))
breast_cancer1
```


```{r}
#pairs(breast_cancer1[,3:22])
cor(breast_cancer1[,3:23])
summary(breast_cancer1[,3:23])
```
# From the correlation, we can notice that variables:[ radius_mean,perimeter_mean,area_mean,concavity_mean,concave_points_mean,radius_worst,perimeter_worst,area_worst,concavity_worst,concave_points_worst] are relatively more correlated to "isM".
# The mean and worst values of most variables are highly correlated.


#fit full model
```{r}
breast_cancer1 = breast_cancer1 %>% select(-id_number, -diagnosis)
breast_cancer1
```

```{r}
model.null = glm(isM ~ 1,
                 data=breast_cancer1,
                 family = binomial(link="logit")
                 )
```

```{r}
model.full = glm(isM ~ .,
                 data=breast_cancer1,
                 family = binomial(link="logit")
                 )
```

```{r}
step(model.null,
     scope = list(upper=model.full),
             direction="both",
             test="Chisq",
             data=breast_cancer1)
```



# Best Multiple logistic model
```{r}
model.final = glm(isM ~ smoothness_worst + texture_worst + concave_points_mean + area_worst + perimeter_mean + symmetry_worst + compactness_mean,data=breast_cancer1,family = binomial(link="logit")
                 )
summary(model.final)
```


Test model accuracy
```{r}
set.seed(1)
train = as_tibble(breast_cancer1[sample(nrow(breast_cancer1), nrow(breast_cancer1)*.7),])
test = anti_join(breast_cancer1, train, by = NULL)
```

```{r}
mod = glm(isM ~ smoothness_worst + texture_worst + concave_points_mean + area_worst + perimeter_mean + symmetry_worst + compactness_mean, data=train,family = binomial(link="logit")
                 )
```

```{r}
test$model_prob <- predict(mod, test, type = "response")
test <- test  %>% mutate(pred = 1*(model_prob > .53) + 0,
                                 isM)
test <- test %>% mutate(accurate = 1*(pred == isM))
sum(test$accurate)/nrow(test)
```







#Full data with SE



```{r}
breast_cancer2 <- breast_cancer
breast_cancer2$isM <- as.factor(ifelse(breast_cancer$diagnosis=="M",1,0))
breast_cancer2$isM <- as.numeric(as.character(breast_cancer2$isM))
breast_cancer2 = breast_cancer2 %>% select(-id_number, -diagnosis)
breast_cancer2
```

```{r}
model.null2 = glm(isM ~ 1,
                 data=breast_cancer2,
                 family = binomial(link="logit")
                 )
```

```{r}
model.full2 = glm(isM ~ .,
                 data=breast_cancer2,
                 family = binomial(link="logit")
                 )
```

```{r}
step(model.null2,
     scope = list(upper=model.full2),
             direction="both",
             test="Chisq",
             data=breast_cancer2)
```

```{r}
model.final2 = glm(formula = isM ~ smoothness_worst + texture_worst + symmetry_worst + 
    compactness_se + concavity_mean + texture_se + area_se + 
    concave_points_worst + area_worst, family = binomial(link = "logit"), 
    data = breast_cancer2)
summary(model.final2)
```

```{r}
set.seed(1)
train = as_tibble(breast_cancer2[sample(nrow(breast_cancer2), nrow(breast_cancer2)*.7),])
test = anti_join(breast_cancer2, train, by = NULL)
```

```{r}
mod = glm(isM ~ smoothness_worst + texture_worst + symmetry_worst + 
    compactness_se + concavity_mean + texture_se + area_se + 
    concave_points_worst + area_worst, data=train,family = binomial(link="logit")
                 )
```

```{r}
test$model_prob <- predict(mod, test, type = "response")
test <- test  %>% mutate(pred = 1*(model_prob > .53) + 0,
                                 isM)
test <- test %>% mutate(accurate = 1*(pred == isM))
sum(test$accurate)/nrow(test)
```




# model with only mean


```{r}
breast_cancerM <- breast_cancer[-c(13:32)]
breast_cancerM$isM <- as.factor(ifelse(breast_cancer$diagnosis=="M",1,0))
breast_cancerM$isM <- as.numeric(as.character(breast_cancerM$isM))
breast_cancerM = breast_cancerM %>% select(-id_number, -diagnosis)
breast_cancerM
```

```{r}
model.nullM = glm(isM ~ 1,
                 data=breast_cancerM,
                 family = binomial(link="logit")
                 )
```

```{r}
model.fullM = glm(isM ~ .,
                 data=breast_cancerM,
                 family = binomial(link="logit")
                 )
```

```{r}
step(model.nullM,
     scope = list(upper=model.fullM),
             direction="both",
             test="Chisq",
             data=breast_cancerM)
```

```{r}
model.finalM = glm(formula = isM ~ concave_points_mean + texture_mean + area_mean + 
    perimeter_mean + smoothness_mean + symmetry_mean, family = binomial(link = "logit"), 
    data = breast_cancerM)
summary(model.finalM)
```

```{r}
set.seed(1)
train = as_tibble(breast_cancerM[sample(nrow(breast_cancerM), nrow(breast_cancerM)*.7),])
test = anti_join(breast_cancerM, train, by = NULL)
```

```{r}
mod = glm(isM ~ concave_points_mean + texture_mean + area_mean + 
    perimeter_mean + smoothness_mean + symmetry_mean, data=train,family = binomial(link="logit")
                 )
```

```{r}
test$model_prob <- predict(mod, test, type = "response")
test <- test  %>% mutate(pred = 1*(model_prob > .53) + 0,
                                 isM)
test <- test %>% mutate(accurate = 1*(pred == isM))
sum(test$accurate)/nrow(test)
```





# model with only SE


```{r}
breast_cancerS <- breast_cancer[c(1,2,13:22)]
breast_cancerS$isM <- as.factor(ifelse(breast_cancer$diagnosis=="M",1,0))
breast_cancerS$isM <- as.numeric(as.character(breast_cancerS$isM))
breast_cancerS = breast_cancerS %>% select(-id_number, -diagnosis)
breast_cancerS
```

```{r}
model.nullS = glm(isM ~ 1,
                 data=breast_cancerS,
                 family = binomial(link="logit")
                 )
```

```{r}
model.fullS = glm(isM ~ .,
                 data=breast_cancerS,
                 family = binomial(link="logit")
                 )
```

```{r}
step(model.nullS,
     scope = list(upper=model.fullS),
             direction="both",
             test="Chisq",
             data=breast_cancerS)
```

```{r}
model.finalS = glm(formula = isM ~ area_se + radius_se + compactness_se + fractal_dimension_se, family = binomial(link = "logit"), 
    data = breast_cancerS)
summary(model.finalS)
```

```{r}
set.seed(1)
train = as_tibble(breast_cancerS[sample(nrow(breast_cancerS), nrow(breast_cancerS)*.7),])
test = anti_join(breast_cancerS, train, by = NULL)
```

```{r}
mod = glm(isM  ~ area_se + radius_se + compactness_se + fractal_dimension_se, data=train,family = binomial(link="logit")
                 )
```

```{r}
test$model_prob <- predict(mod, test, type = "response")
test <- test  %>% mutate(pred = 1*(model_prob > .53) + 0,
                                 isM)
test <- test %>% mutate(accurate = 1*(pred == isM))
sum(test$accurate)/nrow(test)
```



# model with only worst


```{r}
breast_cancerW <- breast_cancer[c(1,2,23:32)]
breast_cancerW$isM <- as.factor(ifelse(breast_cancer$diagnosis=="M",1,0))
breast_cancerW$isM <- as.numeric(as.character(breast_cancerW$isM))
breast_cancerW = breast_cancerW %>% select(-id_number, -diagnosis)
breast_cancerW
```

```{r}
model.nullW = glm(isM ~ 1,
                 data=breast_cancerW,
                 family = binomial(link="logit")
                 )
```

```{r}
model.fullW = glm(isM ~ .,
                 data=breast_cancerW,
                 family = binomial(link="logit")
                 )
```

```{r}
step(model.nullW,
     scope = list(upper=model.fullW),
             direction="both",
             test="Chisq",
             data=breast_cancerW)
```

```{r}
model.finalW = glm(formula = isM ~ smoothness_worst + texture_worst + concave_points_worst +
    area_worst, family = binomial(link = "logit"), 
    data = breast_cancerW)
summary(model.finalW)
```

```{r}
set.seed(1)
train = as_tibble(breast_cancerW[sample(nrow(breast_cancerW), nrow(breast_cancerW)*.7),])
test = anti_join(breast_cancerW, train, by = NULL)
```

```{r}
mod = glm(formula = isM ~ smoothness_worst + texture_worst + concave_points_worst +
    area_worst, family = binomial(link = "logit"), 
    data = train)
```

```{r}
test$model_prob <- predict(mod, test, type = "response")
test <- test  %>% mutate(pred = 1*(model_prob > .53) + 0,
                                 isM)
test <- test %>% mutate(accurate = 1*(pred == isM))
sum(test$accurate)/nrow(test)
```



